Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-08-22 paper/article_jos.qmd
vf

---
title: "Improving real-time trend estimates using local parametrisation of polynomial regression filters"
format:
  docx:
    toc: false
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    keep-tex: true
    toc: false
    fig-pos: 'H'
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
    cite-method: biblatex
crossref:
  tbl-prefix: table
  fig-prefix: figure
  sec-prefix: section
  eq-prefix: equation

@@

@@

To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $y_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
The moving average $M_{\boldsymbol\theta}$ transforms $y_t$ into:
$$
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
$$
The function $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
It can be rewritten as:
$$
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
$$
where $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
We thus have:
\begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*}
The phase shift is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure it in number of periods.
For all symmetric moving averages we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.

@@

<!-- ### Noise reduction -->

<!-- All time series are affected by noise, which can blur the extraction of trends and cycles. -->
<!-- This is why we try to reduce this noise (by reducing its variance) while preserving the relevant trends. -->
<!-- The sum of the squares of the coefficients $\sum_{k=-p}^{+f}\theta_k^2$ is the *reduction of variance* ratio. -->

<!-- Indeed, let $\{\varepsilon_t\}$ be a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$: -->
<!-- $$ -->
<!-- \V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}} -->
<!-- = \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}= -->
<!-- \sigma^2\sum_{k=-p}^{+f} \theta_k^2. -->
<!-- $$ -->

>>>>

---
title: "Improving real-time trend estimates using local parametrisation of polynomial regression filters"
format:
  docx:
    toc: false
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    keep-tex: true
    toc: false
    fig-pos: 'H'
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
    cite-method: biblatex
    output-file: main_document
crossref:
  tbl-prefix: table
  fig-prefix: figure
  sec-prefix: section
  eq-prefix: equation

@@

::: {.content-visible when-format="docx"}
$$
\DeclareMathOperator{\e}{e}
\newcommand\transp[1]{{#1}^T}
\newcommand\1{{\mathbb 1}}
\newcommand{\E}[1]{\mathbb{E}\left[#1\right]}
\newcommand\ud{\,\mathrm{d}}
$$
:::

@@

To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $y_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
The moving average $M_{\boldsymbol\theta}$ transforms $y_t$ into:
$$
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
$$
The function $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{2\pi i \omega k}$.
It can be rewritten as:
$$
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
$$
where $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
We thus have:
\begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*}
The phase shift is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure it in number of periods.
For all symmetric moving averages we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.

@@

### 2024-08-22 paper/main_document.tex
vf

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(y_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(y_t\) into: \[
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can alternatively be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
or
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. We thus have: \begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*} The phase shift is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to measure it in number of periods. For all symmetric moving averages we
have \(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).

>>>>

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(y_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(y_t\) into: \[
Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can alternatively be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{i \omega k}\)
or
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. We thus have: \begin{align*}
\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
\end{align*} The phase shift is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to measure it in number of periods. For all symmetric moving averages we
have \(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).
